### 2020-03-03 doc/icfp20/counterpoint.tex
proof read Section 3.3; add discussion on Haskell '17

\citet{cong-cp} implement correct-by-construction counterpoint
using an older version of Music Tools.
The idea is to represent counterpoint as a list-like datatype,
where the base cases enforce a valid ending and the inductive case
guarantees correct uses of intervals and motion (by means of
implicit arguments representing proofs).

Our representation of counterpoint improves on the existing
implementation in two ways.
First, while \citet{cong-cp} rely on the Agda type checker to report
errors in the counterpoint, we use our own type checker (implemented
as the \texttt{checkXXX} functions) and type errors (defined as the
\texttt{XXXError} datatypes).
This allows us to produce error messages that clearly state what
kind of mistake the programmer has made.
Suppose we have replaced the second interval of Figure~\ref{fig:fs} by
an octave (8th) of D, breaking the rule that perfect intervals cannot
be approached by parallel motion.
In our implementation, this results in an error in the
\texttt{motionOk} field, which requires a proof of
\texttt{checkMotion (firstBar :: middleBars) \(\equiv\) []}:

@@

\noindent In \citet{cong-cp}'s implementation, on the other hand,
the error results in the first \texttt{PitchInterval} being
highlighted in yellow, with the following error message displayed
in the Agda buffer:

\begin{alltt}
\_13 : motionOk (c 5 , c 6) (d 5 , d 6)
\end{alltt}

\noindent This indicates the existence of an unsolved metavariable.
To a non-expert user, it may be difficult to connect this message to
the musical error.

The second improvement from \citet{cong-cp} is that, instead of
incorporating all the rules into the counterpoint datatype, we define
each of them as a separate function.
This makes it possible to reuse certain rules in a different context,
as we will see in the next section.

>>>>

The type-theoretical formalization of counterpoint has previously
been attempted by \citet{cong-cp}.
They represent correct counterpoint as an Agda datatype akin to
a list, where the base cases enforce a valid ending and the inductive
case guarantees correct uses of intervals and motion (by means of
implicit arguments representing proofs).

Our representation of counterpoint improves on the existing
implementation in two ways.
First, while \citet{cong-cp} rely on the Agda type checker to report
errors in the counterpoint, we use our own type checker (implemented
as the \texttt{checkXXX} functions) and type errors (defined as the
\texttt{XXXError} datatypes).
This allows us to produce error messages that clearly state what
kind of mistake the programmer has made.
Suppose we have replaced the second interval of Figure~\ref{fig:fs} by
an octave (8th) of D, breaking the rule that perfect intervals cannot
be approached by parallel motion.
In our implementation, this causes an error in the \texttt{motionOk}
field, with the following message:

@@

\noindent In \citet{cong-cp}'s implementation, on the other hand,
the invalid motion results in an unsolved metavariable \texttt{\_13},
representing the proof that the two intervals form a valid motion:

\begin{alltt}
\_13 : motionOk (c 5 , per8) (d 5 , per8)
\end{alltt}

\noindent Compared to the error message produced by our
implementation, this lacks the exact information about the motion
between the two intervals.

We would also like to note that the attempt to precisely reporting
musical errors has appeared in the work by \citet{szamozvancev-welltyped}.
They formalize various musical rules, including those for intervals
and motion, using GHC's custom type error feature.
However, they do not aim to fully support species counterpoint,
and thus do not cover as many rules as we do.

The second improvement from \citet{cong-cp} is that, instead of
incorporating all the rules into a counterpoint datatype, we define
each of them as a separate function and combine them using a
record type.
This makes it easy to switch between strict and relaxed rule sets:
we just need to add or remove certain fields.
It also allows us to reuse some of the rules in a context other than
counterpoint, as we will see in the next section.

### 2020-03-03 doc/icfp20/music.tex
add figure desciption and remove CCSXML
proof read Section 3.3; add discussion on Haskell '17

\newcommand{\Interval}{
  \begin{figure}
  \includegraphics[width=12cm]{fig/interval.png} \\
  \begin{flushleft}
    \begin{footnotesize}
      \hspace{1.45cm} \texttt{per1 \hspace{0.5mm} min2  \hspace{2.5mm}
        maj2 \hspace{1.2mm} min3 \hspace{0.5mm} maj3 \ per4
        \hspace{1mm}aug4
        \hspace{0.5mm}  per5\hspace{1.2mm}  min6\hspace{1.6mm} maj6
        \hspace{1.6mm}min7\hspace{1.2mm} maj7\hspace{1.2mm} per8} \\
      consonant? \hspace{1.5mm} yes \hspace{4.2mm} no \hspace{6.5mm}  no
      \hspace{4.7mm} yes \hspace{3.8mm}  yes \hspace{3.7mm} no \hspace{3.5mm}
      no \hspace{4.3mm} yes \hspace{3.8mm} yes \hspace{3.2mm} yes
      \hspace{3.7mm} no \hspace{4mm} no \hspace{3.8mm} yes
    \end{footnotesize}
  \end{flushleft}
  \caption{13 Kinds of Intervals}
  \label{fig:interval}
 \end{figure}
}

@@

An \emph{interval} represents the difference in pitch between
two notes.
There are 13 kinds of interval within an octave
(Figure~\ref{fig:interval}, and these intervals can be classified from
several different perspectives:
(i) major or minor; (ii) consonant or dissonant; and (iii) perfect
or imperfect.
We define intervals as a datatype \texttt{Interval}, where each
constructor represents one of the 13 intervals.

>>>>

\newcommand{\Interval}{
  \begin{figure}
  \includegraphics[width=12cm]{fig/interval.png} \\
  \begin{flushleft}
    \begin{footnotesize}
      \hspace{1.45cm} \texttt{per1 \hspace{0.5mm} min2  \hspace{2.5mm}
        maj2 \hspace{1.2mm} min3 \hspace{0.5mm} maj3 \ per4
        \hspace{1mm}aug4
        \hspace{0.5mm}  per5\hspace{1.2mm}  min6\hspace{1.6mm} maj6
        \hspace{1.6mm}min7\hspace{1.2mm} maj7\hspace{1.2mm} per8} \\
      consonant? \hspace{1.5mm} yes \hspace{4.2mm} no \hspace{6.5mm}  no
      \hspace{4.7mm} yes \hspace{3.8mm}  yes \hspace{3.7mm} no \hspace{3.5mm}
      no \hspace{4.3mm} yes \hspace{3.8mm} yes \hspace{3.2mm} yes
      \hspace{3.7mm} no \hspace{4mm} no \hspace{3.8mm} yes
    \end{footnotesize}
  \end{flushleft}
  \caption{13 Kinds of Intervals}
  \Description{13 Kinds of Intervals}
  \label{fig:interval}
 \end{figure}
}

@@

An \emph{interval} represents the difference in pitch between
two notes.
There are 13 kinds of interval within an octave
(Figure~\ref{fig:interval}), and these intervals can be classified from
several different perspectives:
(i) major or minor; (ii) consonant or dissonant; and (iii) perfect
or imperfect.
We define intervals as a datatype \texttt{Interval}, where each
constructor represents one of the 13 intervals.
